---
title: "Homework 3"
author: "Tadeh Peroomian"
date: "2026-05-26"
format: pdf
editor: source
execute:
  warning: false
---

# Homework Setup

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(here)


kelp <- read.csv("data/temp_kelp.csv")
caffeine <- read.csv("data/caffeinated_drinks.csv")

```

# Problem 1

# a.

The two appropriate tests are the Pearson correlation and the Spearman rank correlation. Pearson correlation measures the strength of a linear relationship between two continuous variables and assumes the data are normally distributed, while Spearman rank correlation measures the strength of a monotonic relationship using ranked data and does not require normality.

# b.

```{r}
#| label: visualization
#base layer: ggplot
ggplot(data = kelp,
       #x-axis: temperature
       #y-axis: kelp elongation
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
  #second layer: points for scatterplot
geom_point(color = "green", 
           size = 3) +
  #relabelling axes and title
  labs(x = "Temperature (°C)",
       y = "Kelp Elongation Rate (cm/day)",
       title = "As Temperature Increases, Kelp Frond Elongation Rate Decreases"
  ) +
  #changing theme from base
  theme_bw()

```


# c.

```{r}
#| label: assumptions
#running spearman rank correlation with temp and kelp elongation
cor.test(kelp$temp_c,
         kelp$kelp_elong,
         method = "spearman"
)

```

Spearman's p = -0.69

For the Spearman Rank Correlation, I checked for a monotonic relationship. I used the scatterplot in part 1b, which suggested a monotonic relationship, meaning that the assumption for Spearman's was met.

# d.

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Spearman Rank Correlation, because it does not require the relationship between temperature and kelp elongation to be linear or the data to be normally distributed. We found a moderate to strong relationship between temperature and kelp frond elongation rate(Spearman's ρ = -0.69, S = 9216.1, p < 0.001, ⍺ = 0.05).

# e.

The Spearman Rank Correlation test shows whether there is a consistent monotonic relationship between temperature and giant kelp frond elongation rate. The Spearman's ρ = -0.69 suggests a moderate to strong negative relationship between temperature and kelp frond elongation rate. Thus, kelp frond elongates at a higher rate at lower temperatures, and inversely at a lower rate at higher temperatures.

# f.

```{r}
#| label: double-check
#running pearson rank correlation with temp and kelp elongation
cor.test(kelp$temp_c,
         kelp$kelp_elong,
         method = "pearson"
)
```

The Pearson Correlation test provided me with Pearson's r = -0.69, which is the same as the Spearman's ρ from Problem 1c. Pearson correlation tests for a linear relationship using raw values of temperature and kelp elongation rate, while Spearman correlation tests for a monotonic relationship using ranked values of the same variables. Both tests led me to the same result that there is a moderate to strong negative relationship between temperature and kelp frond elongation rate.


# Problem 2

# a.

```{r}
#| label: caffeinated-drinks-bar
#removing uppercase and spaces/periods
clean_caffeine <- caffeine |> 
  clean_names() 
#creating number of drinks data point
drink_counts <- clean_caffeine |>
  count(time_of_day, name = "num_drinks") 
#base layer: ggplot
ggplot(drink_counts, 
       #x-axis: time of day
       #y-axis: number of drinks
       #filling color based on time of day
       aes(x = time_of_day, 
           y = num_drinks, 
           fill = time_of_day)) +
  #second layer: bar plot
  geom_col() +
  #relabeling title and axes
  labs(title = "What time of Day are more drinks consumed?",
    x = "Time of Day",
    y = "Number of Drinks (count)"
  ) +
  #recoloring based on Time of Day
  scale_fill_manual(values = c("Morning" = "skyblue", 
                               "Afternoon" = "tomato",
                               "Evening" = "limegreen")) +
  #changing theme
  theme_bw() +
  #removing legend
  theme(legend.position = "none")

```

```{r}
#| label: caffeinated-drinks-line
#mutating times on the data table
drink_data <- clean_caffeine |> 
  mutate(hour = as.numeric(format(
    as.POSIXct(drink_date_and_time, format = "%m/%d/%Y %I:%M %p"),
    "%H"
  )))

#creating drinks per hour count
hourly_counts <- drink_data |>
  count(hour, name = "num_drinks")

#base layer: ggplot
ggplot(hourly_counts, aes(x = hour, y = num_drinks)) +
    #second layer: points
  geom_point(color = "purple", size = 3) +
    #relabeling title and axes
  labs(title = "Drink Consumption Tends to Increase Later in the Day",
       x = "Hour of Day (24-hour time)",
       y = "Number of Drinks (count)") +
   #changing theme from base
  theme_bw()
```





