---
title: "Homework 3"
author: "Tadeh Peroomian"
date: "2026-05-26"
format: pdf
editor: source
execute:
  warning: false
---

#. Homework Setup

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(here)


kelp <- read.csv("data/temp_kelp.csv")
caffeine <- read.csv("data/caffeinated_drinks.csv")

```

# Problem 1

#. a

The two appropriate tests are the Pearson correlation and the Spearman rank correlation. Pearson correlation measures the strength of a linear relationship between two continuous variables and assumes the data are normally distributed, while Spearman rank correlation measures the strength of a monotonic relationship using ranked data and does not require normality.

#. b

```{r}
#| label: visualization
#base layer: ggplot
ggplot(data = kelp,
       #x-axis: temperature
       #y-axis: kelp elongation
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
  #second layer: points for scatterplot
geom_point(color = "green", 
           size = 3) +
  #relabelling axes and title
  labs(x = "Temperature (°C)",
       y = "Kelp Elongation Rate (cm/day)",
       title = "As Temperature Increases, Kelp Frond Elongation Rate Decreases"
  ) +
  #changing theme from base
  theme_bw()

```


#. c

```{r}
#| label: assumptions
#running spearman rank correlation with temp and kelp elongation
cor.test(kelp$temp_c,
         kelp$kelp_elong,
         method = "spearman"
)

```

Spearman's p = -0.69

For the Spearman Rank Correlation, I checked for a monotonic relationship. I used the scatterplot in part 1b, which suggested a monotonic relationship, meaning that the assumption for Spearman's was met.

#. d

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Spearman Rank Correlation, because it does not require the relationship between temperature and kelp elongation to be linear or the data to be normally distributed. We found a moderate to strong relationship between temperature and kelp frond elongation rate(Spearman ρ = -0.69, S = 9216.1, p < 0.001, ⍺ = 0.05).

#. e

The Spearman Rank Correlation test shows whether there is a consistent monotonic relationship between temperature and giant kelp frond elongation rate. The Spearman's ρ = -0.69 suggests a moderate to strong negative relationship between temperature and kelp frond elongation rate. Thus, kelp frond elongates at a higher rate at lower temperatures, and inversely at a lower rate at higher temperatures.










