---
title: "Homework 3"
author: "Tadeh Peroomian"
date: "2026-05-26"
format: pdf
editor: source
---

#. Homework Setup

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(here)


kelp <- read.csv("data/temp_kelp.csv")
caffeine <- read.csv("data/caffeinated_drinks.csv")

```

# Problem 1

#. a

The two appropriate tests are the Pearson correlation and the Spearman rank correlation. Pearson correlation measures the strength of a linear relationship between two continuous variables and assumes the data are normally distributed, while Spearman rank correlation measures the strength of a monotonic relationship using ranked data and does not require normality.

#. b

```{r}
#| label: visualization

ggplot(data = kelp,
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
geom_point(color = "green", 
           size = 3) +
  labs(x = "Temperature (°C)",
       y = "Kelp Elongation Rate (cm/day)",
       title = "As Temperature Increases, Kelp Frond Elongation Rate Decreases"
  ) +
  theme_bw()

```






